Demers Cartogram with Rivers

### 2023-07-06 paper/GHRVis.tex
draft: IV

\documentclass{egpubl}
\usepackage{VMV2023}

% --- for  Annual CONFERENCE
% \ConferenceSubmission   % uncomment for Conference submission
% \ConferencePaper        % uncomment for (final) Conference Paper
% \STAR                   % uncomment for STAR contribution
% \Tutorial               % uncomment for Tutorial contribution
% \ShortPresentation      % uncomment for (final) Short Conference Presentation
% \Areas                  % uncomment for Areas contribution
% \MedicalPrize           % uncomment for Medical Prize contribution
% \Education              % uncomment for Education contribution
% \Poster                 % uncomment for Poster contribution
% \DC                     % uncomment for Doctoral Consortium
%
% --- for  CGF Journal
% \JournalSubmission    % uncomment for submission to Computer Graphics Forum
% \JournalPaper         % uncomment for final version of Journal Paper
%
% --- for  CGF Journal: special issue
% \SpecialIssueSubmission    % uncomment for submission to , special issue
% \SpecialIssuePaper         % uncomment for final version of Computer Graphics Forum, special issue
%                          % EuroVis, SGP, Rendering, PG
% --- for  EG Workshop Proceedings
\WsSubmission      % uncomment for submission to EG Workshop
% \WsPaper           % uncomment for final version of EG Workshop contribution
% \WsSubmissionJoint % for joint events, for example ICAT-EGVE
% \WsPaperJoint      % for joint events, for example ICAT-EGVE
% \Expressive        % for SBIM, CAe, NPAR
% \DigitalHeritagePaper
% \PaperL2P          % for events EG only asks for License to Publish

% --- for EuroVis 
% for full papers use \SpecialIssuePaper
% \STAREurovis   % for EuroVis additional material 
% \EuroVisPoster % for EuroVis additional material 
% \EuroVisShort  % for EuroVis additional material

% !! *please* don't change anything above
% !! unless you REALLY know what you are doing
% ------------------------------------------------------------------------

\usepackage[T1]{fontenc}
\usepackage{dfadobe}

% \usepackage{cite}  % comment out for biblatex with backend=biber
% ---------------------------
\biberVersion
\BibtexOrBiblatex
\usepackage[backend=biber,bibstyle=EG,citestyle=alphabetic,backref=true,url=false]{biblatex}
%\addbibresource{egbibsample.bib}

\renewbibmacro*{doi+eprint+url}{%
    \printfield{doi}%
    \newunit\newblock%
    \iftoggle{bbx:eprint}{%
        \usebibmacro{eprint}%
    }{}%
    \newunit\newblock%
    \iffieldundef{doi}{%
        \usebibmacro{url+urldate}}%
        {}%
    }

\addbibresource{references.bib}

\newcommand{\citetitlea}[1]{\citefield{#1}{title}}

% ---------------------------  
\electronicVersion
\PrintedOrElectronic
% for including postscript figures
% mind: package option 'draft' will replace PS figure by a filename within a frame
\ifpdf \usepackage[pdftex]{graphicx} \pdfcompresslevel=9
\else \usepackage[dvips]{graphicx} \fi

\usepackage{egweblnk}
% end of prologue

\usepackage{float,hhline,hyperref,array,tabulary,tabularx,etoolbox,multirow}

@@

\usepackage{dblfloatfix}

@@

\usepackage{multicol}

% For cover's captionof
\usepackage[font=normalfont,labelfont=bf]{caption}

\include{preamble/preamble.basic}
\include{preamble/preamble.main}
\include{preamble/preamble.math}
\include{preamble/preamble.draft}

@@

@@

% \author[Q. Wang, K. Xu \& R. S. Laramee]
\author[Authors]
{
% \parbox{\textwidth}{\centering Q. Wang, K. Xu and R. S. Laramee}
% \\
% \parbox{\textwidth}{\centering School of Computer Science, University of Nottingham, UK \\ \begin{center}
\parbox{\textwidth}{\centering Authors }
\\
\parbox{\textwidth}{\centering Affiliation \\ \begin{center}
        \vspace{2em}
        \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
        \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}
    \end{center}
}
}

@@

\pagestyle{plain}

@@

% \section{Acknowledgments}
% This work is funded by the grant EP/S010238/2 from the Engineering and Physical Sciences Research Council (EPSRC). EPSRC is a British Research Council that provides government funding for grants to undertake research and postgraduate degrees in engineering and the physical sciences.

% reset \section command for section title "References"
\let\section=\origsection
\printbibliography

\clearpage

>>>>

\documentclass[Afour,sagev,times]{sagej}

\usepackage{moreverb,url}

\usepackage{float,hhline,array,tabulary,tabularx,etoolbox,multirow, multicol}

@@

@@

\include{preamble/preamble.main}
\include{preamble/preamble.math}

\usepackage[colorlinks,citecolor=red,urlcolor=red]{hyperref}

\newcommand\BibTeX{{\rmfamily B\kern-.05em \textsc{i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\setcounter{secnumdepth}{3}

@@

@@

% \includegraphics[width=\textwidth,keepaspectratio]{figure/cover.png}
%         \captionof{figure}{The River Thames passing through a cartographic representation of the NHS CCGs in London and surrounding areas.}

\keywords{Cartogram, Demers Cartogram}

@@

% \bibliographystyle{mslapa}
\urlstyle{same}
\bibliographystyle{plainnat}
\bibliography{references}

\cleardoublepage

### 2023-07-06 paper/section/evaluation.tex
draft: IV

\newcommand{\pCount}{24 }
\section{User-Centered Evaluation}

@@

\subsection{Study Hypothesis}

@@

\subsection{User Study Variables}

@@

\subsection{User Study Design}
In this section we describe the user study participants, datasets, and the experimental procedure.

@@

\begin{itemize}
    \item Gender: 10 females and 14 males
    \item Age Group: 18-24 (16), 25-29 (8)
    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
\end{itemize}

@@

\begin{itemize}
    \item Population
    \item Under 75 mortality from cardiovascular disease
    \item Emergency admissions for alcohol-related liver disease
    \item Alcohol-specific admission and readmission
\end{itemize}

@@

\subsubsection{Procedure}

@@

\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos. The instructions are designed to help participants understand the concepts used in the tasks. Instructional videos are available at \URL{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.

@@

The materials for P1-P4 are available on the Open Science Foundation website at \URL{https://osf.io/q39w7}.

\subsection{User Study Results}

>>>>

\newcommand{\pCount}{24 }
\section{User-Centered Evaluation}\label{sec:{User-Centered Evaluation}}

@@

\subsection{Study Hypothesis}\label{subsec:{Study Hypothesis}}

@@

\subsection{User Study Variables}\label{subsec:{User Study Variables}}

@@

\subsection{User Study Design}\label{subsec:{User Study Design}}

In this section we describe the user study participants, datasets, and the experimental procedure.

@@

\begin{itemize}
    \setlength\itemsep{0px}
    \item Gender: 10 females and 14 males
    \item Age Group: 18-24 (16), 25-29 (8)
    \item Education: 1 PhD, 11 Master's, 8 Bachelor's, 4 Others
\end{itemize}

@@

\begin{itemize}
    \setlength\itemsep{0px}
    \item Population
    \item Under 75 mortality from cardiovascular disease
    \item Emergency admissions for alcohol-related liver disease
    \item Alcohol-specific admission and readmission
\end{itemize}

@@

\subsubsection{Procedure}
\label{subsec:{Procedure}}

@@

\textbf{P1:} The participants are asked to listen to instructions and training provided as both text and videos. The instructions are designed to help participants understand the concepts used in the tasks. Instructional videos are available at \url{https://www.youtu.be/playlist?list=PLL7sHvxLtD75fMtrUQrAdddjt3wfFkcWz}.

@@

The materials for P1-P4 are available on the Open Science Foundation website at \url{https://osf.io/q39w7}.

\subsection{User Study Results}
\label{subsec:{User Study Results}}

### 2023-07-06 paper/section/introduction.tex
draft: adjust layout
draft: IV

Among the four types of cartogram categorized in a survey by \citea{nusrat2016State} (See \Cref{sec:{Related Work}} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes. Also, the comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions. 
We hypothesize that introducing rivers improves the overall legibility of a cartogram. 
By \textit{legibility} we mean readability and ability to interpret the cartogram.  
We test this by asking user-study participants to perform correspondence and location tasks.
To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
We present a user study that demonstrates its effectiveness.

@@

\begin{itemize}
    \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,
    \item A novel hybrid layout algorithm that combines node positions with features such as rivers,
    \item A user study evaluation of the technique with an application to EHRs.
\end{itemize}

>>>>

Among the four types of cartogram categorized in a survey by \citet{nusrat2016State} (See \Cref{sec:RelatedWork} for definitions of contiguous, non-contiguous, rectangular, and Dorling), a trade-off is made between types of accuracy (See \Cref{table:accuracy}).
For this project, we focus on non-contiguous cartograms like Demers cartograms because they facilitate statistical comparison between regions, they can make good use of screen space, and comparison of regions is useful when studying Electronic Health Records (EHR) data.
Demers cartograms offer the advantage in cases where the data is not directly correlated to region sizes. Also, the comparison of magnitudes becomes an area estimation task, which is more effective for a numeric data encoding \cite{munzner2014Visualization}.
Building on Demers cartograms \cite{ian2002Cartogram}, we introduce and develop novel features, such as rivers, aiming to improve the readability and geographical accuracy without sacrificing statistical accuracy.
Standard Demers cartograms are composed of square nodes. As such, this can reduce their legibility.
We implement a new hybrid cartographic layout algorithm that combines rivers with the placement of the nodes representing geo-referenced regions. 
We hypothesize that introducing rivers improves the overall legibility of a cartogram. 
By \textit{legibility} we mean readability and ability to interpret the cartogram.  
To assess this hypothesis, we designed an experimental setup where participants engaged in correspondence and location tasks as part of a user study.
To reduce error and make efficient use of screen space, the algorithm also updates the position of rivers to accommodate the node layout.
We then apply the algorithm to a real-world case study using EHR data to evaluate the result.
We present a user study that demonstrates its effectiveness.

@@

\begin{itemize}
    \setlength\itemsep{0px}
    \item A new variant of Demers cartograms that incorporates rivers to improve readability and recognizability,
    \item A novel hybrid layout algorithm that combines node positions with features such as rivers,
    \item A user study evaluation of the technique with an application to EHRs.
\end{itemize}

### 2023-07-06 paper/section/main.tex
draft: IV

\section{Demers Cartogram with Rivers}
 {
  \begin{figure*}[tb!]
      \centering
      \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
      \caption{An overview of our hybrid layout algorithm incorporating rivers. See also \Cref{alg:UpdateLayout} in \Cref{sec:{Demers Cartogram with Rivers}} for more detail. % See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
       For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
      \label{fig:flowchart}
  \end{figure*}
 }

@@

\subsection{River Intersection Testing}

@@

\subsection{Translating Rivers}

@@

\subsection{Process Stalemates}

@@

\subsection{Terminating the Algorithm}

The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user. We adopt the maximum cartographic error from \citea{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
$$

@@

\subsection{User Options}

@@

{
\begin{figure*}[tb!]
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio]{figure/UI.png}
    \caption{A screenshot of the user interface. User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers. Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic. See \Cref{subsec:{User Options}} for more on user options. In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
    \label{fig:overview}
\end{figure*}
}

>>>>

\section{Demers Cartogram with Rivers}
\label{sec:main}

 {
  \begin{figure*}[tb!]
      \centering
      \includegraphics[width=0.8\textwidth,height=\textheight,keepaspectratio]{figure/flowchart.png}
      \caption{An overview of our hybrid layout algorithm incorporating rivers. See also \Cref{alg:UpdateLayout} in Section \ref{sec:main} for more detail. % See \Cref{fig:flowchart-stalemate} for the logic of processing a stalemate.
       For illustration purposes, we show the rendered views alongside the logical views representing the actual computation.}
      \label{fig:flowchart}
  \end{figure*}
 }

@@

\subsection{River Intersection Testing}
\label{subsec:{River Intersection Testing}}

@@

\subsection{Translating Rivers}
\label{subsec:{Translating Rivers}}

@@

\subsection{Process Stalemates}
\label{subsec:{Process Stalemates}}

@@

\subsection{Terminating the Algorithm}
\label{subsec:{Terminating the Algorithm}}

The algorithm terminates when $ \nodeCartographicError $ reaches $ \nodeCartographicErrorMax $, the error tolerance set by the user. We adopt the maximum cartographic error from \citet{alam2015Quantitative}, namely: $$ \nodeCartographicErrorMax = \max _{\node \in \nodeList} \frac{|\node_i-\node_{i_t}|}{\max \{\node_i, \node_{i_t}\}}
$$

@@

\subsection{User Options}\label{subsec:{UserOptions}}

@@

{
\begin{figure*}[tb!]
    \centering
    \includegraphics[width=0.7\textwidth,keepaspectratio]{figure/UI.png}
    \caption{A screenshot of the user interface. User options are provided to adjust the terminating conditions (size and error), color mapping, visibility of nodes and rivers. Other options include the ability to control the overlap removal behavior: rivers can be static or dynamic. See \Cref{subsec:{UserOptions}} for more on user options. In this figure, $ \nodeCartographicErrorMax $ = 1.875\%, and an $ \nodeTopologicalError $ of 9.63\% is eliminated.}
    \label{fig:overview}
\end{figure*}
}

### 2023-07-06 paper/section/related.tex
draft: adjust layout
draft: IV

\section{Related Work}

@@

\citea{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological. Each cartogram design may make various types of accuracy trade-offs between dimensions. We provide a comparison of these trade-offs as introduced by \citea{nusrat2016State} in Table \ref{table:accuracy}. In addition, we include the Demers as it is the focus of our work.

@@

			% \citea{auber2007Geographical} & \citetitlea{auber2007Geographical} & US & 19 & 2007 \\
			% \hline
			\citea{warf2008Geography} & \citetitlea{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
			\hline
			\citea{sun2010Effectiveness} & \citetitlea{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
			\hline
			\citea{cruz2017Adapted} & \citetitlea{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
			\hline
			\citea{tong2018Cartograms} & \citetitlea{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
			\hline
			\citea{gao2020Visualising} & \citetitlea{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
			\hline
			\citea{nusrat2020Recognition} & \citetitlea{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
			\hline
			\citea{nickel2022Multicriteria} & \citetitlea{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
			\hline

@@

\thumbnail{dorling}{
	Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
	A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citea{cano2015Mosaic} in their related work section.
	Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
	This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
	This is important in our use case scenario involving EHRs.
	Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}. 
}

@@

\bobgraph{Peer-reviewed Applications: }\citea{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US. \citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US. See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

Our work extends the algorithm described by \citea{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms. Our work enhances that of \citeauthor{tong2018Cartograms} in multiple ways. First, we introduce multiple features (rivers) into the layout, as opposed to a single river. Second, we make topological features dynamic and further improve legibility and geographical accuracy. By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm. In previous work, the river is static and serves merely as a boundary. Third, we improve the algorithm to resolve stalemates. Finally, the way we evaluate the cartograms is also different. \citeauthor{tong2018Cartograms} count river crossings to evaluate error (a statistical metric). Here our focus is on readability, and thus we include a user study.

\bobgraph{Cartograms in Media: }Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news. The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

% \subsection{Contiguous Applications}

% \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.

One of the disadvantages of Dorling and Demers cartograms is legibility. The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret. \citea{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.

>>>>

\section{Related Work}
\label{sec:RelatedWork}

@@

\citet{nusrat2016State} define and summarize three major accuracy dimensions for cartograms: statistical, geographical, and topological. Each cartogram design may make various types of accuracy trade-offs between dimensions. We provide a comparison of these trade-offs as introduced by \citet{nusrat2016State} in \Cref{table:accuracy}. In addition, we include the Demers as it is the focus of our work.

@@

			% \citet{auber2007Geographical} & \citep{auber2007Geographical} & US & 19 & 2007 \\
			% \hline
			\citet{warf2008Geography} & \citep{warf2008Geography} & Dorling & US & 3,142 & \citeyear{warf2008Geography} \\
			\hline
			\citet{sun2010Effectiveness} & \citep{sun2010Effectiveness} & Dorling, Mosaic, Neighbor-preserving & US, China & 34 - 49 & \citeyear{sun2010Effectiveness} \\
			\hline
			\citet{cruz2017Adapted} & \citep{cruz2017Adapted} & Dorling, Non-contiguous, Neighbor-preserving & Portugal & 2,882 & \citeyear{cruz2017Adapted} \\
			\hline
			\citet{tong2018Cartograms} & \citep{tong2018Cartograms} & Demers & England & 209 & \citeyear{tong2018Cartograms} \\
			\hline
			\citet{gao2020Visualising} & \citep{gao2020Visualising} & Dorling & China & 34 & \citeyear{gao2020Visualising} \\
			\hline
			\citet{nusrat2020Recognition} & \citep{nusrat2020Recognition} & Contiguous, Dorling & Portugal & 49 & \citeyear{nusrat2020Recognition} \\
			\hline
			\citet{nickel2022Multicriteria} & \citep{nickel2022Multicriteria} & Contiguous, Demers & US, Netherlands, World & 49 - 342 & \citeyear{nickel2022Multicriteria} \\
			\hline

@@

\thumbnail{dorling}{
	Dorling cartograms as a variant of non-contiguous cartogram generally do not preserve geography and topology.
	A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimensions of interest are represented by the circle size \cite{dorling2011Area}.
	In a Demers cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology, as described by \citet{cano2015Mosaic} in their related work section.
	Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap. Here we focus on Demers cartograms as we use squares to depict regions.
	This style of cartogram offers the advantages that the comparisons between regions are intuitive and screen space utilization is more efficient.
	This is important in our use case scenario involving EHRs.
	Demers cartograms, where regions are represented by squares, often have the advantage of preserving a higher level of topology at the cost of geographical accuracy \cite{ian2002Cartogram}. 
}

@@

\bobgraph{Peer-reviewed Applications: }\citet{warf2008Geography} use a Dorling cartogram to represent religious diversity in the US. \citet{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms. \citet{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citet{tong2018Cartograms} use a Demers cartogram to visualize health-related data by Clinical Commissioning Group (CCG) regions in England, the work introduces a novel technique to remove the overlap of squares based on topological features, aiming to improve both geographical and topological accuracy. \citet{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms using multiple data sets that include demographic, agriculture, and retail data in the US. See \Cref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

Our work extends the algorithm described by \citet{tong2018Cartograms} which incorporates a static topological feature into Demers cartograms. Our work enhances that of \citet{tong2018Cartograms} in multiple ways. First, we introduce multiple features (rivers) into the layout, as opposed to a single river. Second, we make topological features dynamic and further improve legibility and geographical accuracy. By the term \textit{dynamic}, we mean that the position of the rivers is updated as part of the layout algorithm. In previous work, the river is static and serves merely as a boundary. Third, we improve the algorithm to resolve stalemates. Finally, the way we evaluate the cartograms is also different. \citet{tong2018Cartograms} count river crossings to evaluate error (a statistical metric). Here our focus is on readability, and thus we include a user study.

\bobgraph{Cartograms in Media: }Cartograms are a popular choice of representation in covering various topics by the media, with the goal of presenting an unbiased and complete picture of the news. The Washington Post uses cartograms to visualize the US overseas economic assistance, in arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \citet{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citet{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

One of the disadvantages of Dorling and Demers cartograms is legibility. The layout algorithms may displace regions far from their original position and make the maps more difficult to interpret. \citet{nickel2022Multicriteria} present a method to compute stable Demers cartograms with multiple constraints to maintain adjacencies with no overlapping nodes.
